\documentclass[twocolumn]{aastex631}
\input{generated/macros}
\newcommand{\code}[1]{\texttt{#1}}
\shorttitle{A uniform Cat-2 repeater timing census}
\shortauthors{Barbere}
\begin{document}
\title{A Uniform Wait-Time, Periodicity, and Duty-Cycle Census of the
CHIME/FRB Catalog 2 Repeaters}
\author[0009-0008-3289-4447]{Joseph Barbere}
\affiliation{Independent researcher}

\begin{abstract}
CHIME/FRB Catalog 2 \citep{cat2} contains 83 repeating sources, yet its repeater timing has
only been analysed source-by-source: the companion rate census \citep{cook2026} computes
burst rates, DM drifts, and the repeater fraction, but no per-source wait-time clustering
shape, no periodograms, and no duty cycles; all published periodicity results are
single-source campaigns. We compute ONE statistic uniformly across every Cat-2 repeater: the
Weibull wait-time shape $k$ in the \citet{oppermann2018} formalism, a Rayleigh-$Z^2$
periodogram, and an activity-window duty cycle at any significant period. Because the public
exposure release is time-integrated, per-epoch exposure correction is impossible from public
data --- stated plainly --- so periodogram significance comes from a transit-comb-preserving
scramble: each burst keeps its sidereal phase and gets a uniformly redrawn day number, putting
CHIME's daily-visibility aliasing into every null realisation (the Catalog-1 lesson that
exposure-blind FAPs are not rigorous, mechanised). The pipeline re-finds the one in-catalogue
anchor: FRB 20180916B at \fwRealAnchorPeriod\,d ($p=$ \fwRealAnchorP, published 16.35\,d;
\citealt{chime2020period}) with a 90\%-containment duty cycle of \fwRealAnchorDuty. Across the
census (\fwRealN\ sources; \fwRealNStats\ above the 10-burst/30-day completeness cut), the
median Weibull shape is $k=$ \fwRealMedianK, \fwRealNClustered\ sources are clustered at 95\%
confidence, and \fwRealNPeriodic\ beat the scramble null at $p\le0.01$ --- of which only the
anchor has enough cycles across the span to call the peak a period; the others are
activity-epoch degeneracies and we label them as such, not as new periods.
\end{abstract}

\keywords{Radio transient sources (2008) --- Radio bursts (1339) ---
Astronomy data analysis (1858)}

\section{Introduction}
Whether repeating FRB sources share a common temporal character --- Poissonian storms,
long-memory clustering, periodic activity windows --- is a population question, but the
population data have been analysed almost entirely one source at a time. The two known
activity periods (16.35\,d for FRB 20180916B, \citealt{chime2020period}; $\sim$160\,d for
FRB 20121102A, \citealt{rajwade2020}) came from dedicated campaigns; wait-time clustering fits exist for a handful of
loud repeaters \citep[e.g.][]{oppermann2018,frb220912a}; and the Catalog-2 companion census
\citep{cook2026} stops at rates and DM drift, noting that ``repeater burst arrival times are
stochastic; however, they cluster in time'' without modelling it. Catalog 2 \citep{cat2},
with 83 repeaters observed under one instrument, one detection pipeline, and one exposure
accounting for just over five years, is the first dataset on which a homogeneous timing census
is even possible. This paper runs it.

Two scope notes. First, FRB 20240114A and FRB 20240209A --- the two loudest recent repeaters
with claimed long periods --- post-date the Catalog-2 cutoff (2023 September 15) and are not
in the table; the only in-catalogue anchor with an established period is FRB 20180916B, and
re-finding it is the validation gate for everything else. Second, the public exposure product
accompanying the catalog (CANFAR DOI 10.11570/25.0066) is a time-integrated sky map, not a
per-epoch exposure history; we therefore never claim a per-epoch-corrected periodogram, and
build the null differently (\S\ref{sec:method}).

\section{Data and Method}
\label{sec:method}
We use the public burst table (4,539 events; sub-bursts deduplicated to the first component's
\code{mjd\_400} TOA) and its per-source upper-transit exposure hours. Every source flagged
with a \code{repeater\_name} enters the census with a rate row; sources with $\ge$10 bursts
and $\ge$30-day spans (the stated completeness cut) additionally get: (i) a Weibull fit to
their inter-burst wait times with bootstrap CIs on the shape $k$ ($k<1$ with the CI excluding
1 = clustered); (ii) a Rayleigh-$Z^2$ periodogram over 1.5\,d--span/3; (iii) a
90\%-containment activity-window duty cycle at the peak period, quoted only when the peak is
significant.

CHIME is a transit instrument: a source is visible for $\sim$minutes per sidereal day, so
burst TOAs live on a comb of period 0.99727\,d, and any exposure-blind false-alarm estimate
inherits that aliasing. Our null preserves it exactly: each burst keeps its sidereal phase
(its position within the daily transit window) while its day number is redrawn uniformly over
the source's observed span; the periodogram-peak FAP is the fraction of $10^3$ such scrambles
whose maximum $Z^2$ over the same grid reaches the observed peak, $p=(k{+}1)/(n{+}1)$. What
the scramble does NOT carry is day-scale exposure structure the public release omits ---
instrument downtime and per-day exposure gaps are absent from the null (days are redrawn
uniformly), a stated approximation, not a solved problem.

The same transit geometry censors the wait-time data: waits between $\sim$15 minutes and
$\sim$1 sidereal day are unobservable (the source has set), so observed waits are either
intra-window or $\ge$1\,d. The Weibull fit is therefore conditioned on the visibility
function; the missing intermediate waits bias $k$ downward (toward apparent clustering), and
the bias is rate-dependent --- strongest for the highest-rate sources, where intra-window
waits are common (FRB 20220912A, our smallest $k$, is the worst case). We disclose rather
than correct it: the census's $k$ values are comparative, conditioned-on-CHIME-sampling
statistics, not deconvolved intrinsic shapes.
Validation is end-to-end on transit-sampled synthetics: an injected Weibull train ($k=0.4$)
with a 16.35-d, 30\%-duty activity window is recovered in all three statistics, and a Poisson
control with no window stays null (the offline CI leg).

One statistic needs a warning label we attach throughout: a periodogram peak with only a few
cycles across the span is degenerate with a small number of activity epochs. We report
$N_{\rm cyc}={\rm span}/P$ per source and treat peaks with $N_{\rm cyc}\lesssim5$ as
epoch-clustering candidates, not periods, whatever their formal $p$.

\section{Results}
\label{sec:results}
\begin{table*}
\centering
\caption{The census, ranked by burst count (top 20 of \fwRealNStats\ above the completeness
cut; the full machine-readable census ships in \code{results/frbwait\_metrics.json}).
$p=(k{+}1)/(n{+}1)$ against the transit-comb-preserving scramble, $n=999$. Duty cycles
(90\% containment) are quoted only for $p\le0.01$ peaks with $\ge$10 cycles across the span
--- at fewer cycles a ``duty cycle'' would dress up an epoch degeneracy as a period.}
\label{tab:census}
\begin{tabular}{lllllll}
\hline
source & bursts & Weibull $k$ & peak period (d) & cycles & $p$ & duty \\
\hline
\input{generated/census_table}
\end{tabular}
\end{table*}

Table~\ref{tab:census} is the census. The anchor validates the machinery:
FRB 20180916B returns \fwRealAnchorPeriod\,d at the scramble floor ($p=$ \fwRealAnchorP,
107 cycles) against the published 16.35\,d, with duty cycle \fwRealAnchorDuty\ (i.e.
$\sim$3.5\,d at 90\% containment) --- consistent with the published 5-d window containing
\emph{all} bursts and 50\% of them within 0.6\,d \citep{chime2020period}, once containment
conventions are matched. Population-wise: the median Weibull shape among sources above the cut is
\fwRealMedianK, below the Poisson value of 1, and \fwRealNClustered\ sources are individually
clustered at 95\% confidence --- sub-Poissonian clustering is the population norm wherever
the data can measure it, extending the single-source results of \citet{oppermann2018} and
\citet{frb220912a} to a uniform sample. \fwRealNPeriodic\ sources beat the scramble at
$p\le0.01$; besides the anchor, these peaks have $N_{\rm cyc}\lesssim5$ and are
activity-epoch degeneracies under our own warning label --- we explicitly do NOT claim new
periods. Their epochs are real (the same structure drives their small $k$), but
distinguishing a window from a handful of storms needs longer spans.

\section{Discussion and Conclusions}
The census's honest summary: (i) clustering is ubiquitous and quantified uniformly for the
first time at population scale; (ii) exactly one Catalog-2 repeater has a
scramble-significant, many-cycle activity period --- the one already known; (iii) no new
periodicity claims survive the cycles test, and the census quantifies exactly how much data
each source needs before one could. The per-source $k$ distribution is the dataset-level
input population-synthesis and progenitor models have been missing; the machinery (tested,
transit-aware, with the null published alongside) transfers to Catalog 3 unchanged. The main
limitation is stated, not hidden: without a public per-epoch exposure history, duty cycles
inherit the transit-sampling approximation, and quantitative window shapes for
low-declination sources (least daily exposure) are the least constrained.

\textit{Reproducibility.} \code{jansky\_research.frbwait}; every number above is
pipeline-generated (\code{results/frbwait\_metrics.json} $\to$ macros and
Table~\ref{tab:census}). Data: CANFAR DOI 10.11570/25.0066 mirror of the ApJS
machine-readable table \citep{cat2}.

\bibliography{refs}
\bibliographystyle{aasjournal}
\end{document}
