\documentclass[twocolumn]{aastex631}
\input{generated/macros}
\newcommand{\code}[1]{\texttt{#1}}
\shorttitle{A blind reanalysis of the ice-giant radio rotation periods}
\shortauthors{Barbere}
\begin{document}
\title{Why a Blind Modern Reanalysis of the Voyager~2 PRA Data Does Not Recover the
Uranus and Neptune Radio Rotation Periods}
\author[0009-0008-3289-4447]{Joseph Barbere}
\affiliation{Independent researcher}

\begin{abstract}
The rotation periods anchoring the System~III longitude systems of Uranus ($17.24\pm0.01$~h) and
Neptune ($16.11$~h) were fixed by the Voyager~2 Planetary Radio Astronomy (PRA) experiment during
the 1986 and 1989 flybys \citep{warwick1986,warwick1989} and, for the radio data, never reanalysed
with modern statistics --- even as \citet{lamy2025} recently shifted Uranus's period by 28~s using
Hubble UV aurorae. We ask a narrow, answerable question: what do the Voyager radio data yield under a
\emph{blind} modern periodogram, using none of the emission-beaming or magnetic-longitude modelling
of the original determinations? From the open PDS-PPI PRA low-band 6-second encounter volumes
\citep{pdsppi_vg2pra} we build a band-integrated flux series and search it with a Lomb--Scargle
periodogram over a wide, un-tuned 14--20~h window, taking a rotation-block bootstrap and the peak's
spread across frequency sub-bands as the uncertainties. \textbf{The result is a controlled null.} The
pipeline is validated end-to-end: on a synthetic flyby searched in the same wide window it recovers
an injected period (\vgSynInjected~h $\to$ \vgSynRecovered~h) to $\sim$1~min. On the real data it
recovers \emph{neither} planet: the Uranus peak wanders between sub-bands (\vgRealUPeriod~h, sub-band
spread \vgRealUSpread~h, offset $+1.2$~h from $17.24$~h; not recovered, \vgRealURecovers) and the
Neptune peak rails to the search bound (\vgRealNPeriod~h; not recovered, \vgRealNRecovers). Because
the method demonstrably works when a clean rotational signal is present, the real-data failure is a
genuine limitation of blind total-power analysis, not a pipeline artifact: the ice-giant total-power
radio flux is not a clean rotational sinusoid, and recovering the periods requires the geometric
modelling of the original analyses. Even in the best case the flyby precision ceiling is $\sim$1--2~h,
hundreds of times coarser than the modern Hubble result. The contribution is a reproducible,
modern-statistical demonstration of exactly why the historical Voyager radio periods needed the
sophisticated methods they used --- and of how little a naive periodogram of these short flybys can
constrain.
\end{abstract}

\keywords{Planetary science (1255) --- Uranus (1751) --- Neptune (1096) --- Radio astronomy (1338)
--- Planetary magnetospheres (997) --- Time series analysis (1916)}

\section{Introduction}
The rotation period of a giant planet's deep interior is inferred from the periodic modulation of
magnetospheric phenomena locked to the internal magnetic field. For Uranus and Neptune the defining
measurement was the periodicity of auroral radio emission --- Uranus Kilometric Radiation (UKR) and
its Neptune analogue --- recorded by Voyager~2 PRA during the only spacecraft encounters with the ice
giants \citep{warwick1986,warwick1989}. Those periods, $17.24\pm0.01$~h and $16.11$~h, became the
System~III longitude systems still in use.

Two facts motivate a fresh look. First, \citet{lamy2025} used 11~yr of Hubble UV aurora tracking to
refine Uranus's period to $17.247864\pm0.000010$~h --- a 28~s shift and a $10^3$-fold precision gain
over the Voyager value, highlighting that the \emph{radio} value was a single brief-flyby measurement
never revisited with modern statistics. Second, the PRA refurbishing effort \citep{cecconi2017}
reprocessed only Jupiter and Saturn, leaving the ice-giant volumes an unworked archive. The original
period determinations were not simple periodograms: they folded in magnetic-field models and emission
beaming. We therefore ask the complementary, honest question --- how much does a \emph{blind} modern
periodogram of the raw radio flux, using none of that modelling, recover on its own? --- and answer
it, with a validated pipeline, in the negative.

\section{Data}
We use the open PDS-PPI PRA low-band 6-second RDR encounter volumes \citep{pdsppi_vg2pra}, datasets
\code{VG2-U-PRA-3-RDR-LOWBAND-6SEC-V1.0} (Uranus) and \code{VG2-N-PRA-3-RDR-LOWBAND-6SEC-V1.0}
(Neptune). Each is a fixed-width ASCII table of 48-second major frames; a frame carries a
spacecraft-event date and second-of-day plus eight sweeps of one status word and 70 channel
intensities in millibell, the sweeps spaced 6~s apart. The 70 low-band channels run from 1326.0 to
1.2~kHz in 19.2~kHz steps. Expanding the sweeps gives a 6-second-cadence dynamic spectrum spanning
$\sim$\vgRealUSpan~h at Uranus and $\sim$\vgRealNSpan~h at Neptune --- of order 15--30 rotations.

\section{Method}
\label{sec:method}
\textbf{Flux series.} We sum the linear power ($10^{\mathrm{mB}/1000}$ per channel) over the
100--1000~kHz band where the auroral emission dominates, giving one total-power sample per 6-second
sweep.

\textbf{Blind Lomb--Scargle.} A rotation would modulate a continuous, red-noise-dominated emission, so
we search the (log, 0.1-h-binned, quadratically-detrended) flux with a Lomb--Scargle periodogram over
a wide, physically-motivated 14--20~h ice-giant window --- deliberately \emph{not} tuned to bracket
the known values, since a narrow window centred on the answer would manufacture a match. Because the
samples are strongly autocorrelated and enormous in number, the analytic Lomb--Scargle false-alarm
probability is meaningless (absurdly small) and we do not report it. The \emph{statistical}
uncertainty is a rotation-block bootstrap (resample whole rotation-length blocks); the
\emph{systematic} one is the peak's spread across four frequency sub-bands. A coherent signal ---
rotation \emph{or} a viewing-geometry, beaming, or spectral-window modulation --- is achromatic, so a
small sub-band spread confirms a real coherent peak but does \emph{not} establish that it is the
rotation. A discrete-epoch Rayleigh $Z^2_1$ fold \citep{buccheri1983} of detected bursts, reused from
this repository's FRB module, is reported as a weaker cross-check.

\textbf{Recovery criterion.} We call the historical period \emph{recovered} only if the blind peak
lands within one combined-$\sigma$ of it, is band-robust (sub-band spread $<0.6$~h), and is not railed
to a search-window bound --- proximity alone is worthless when $\sigma\sim2$~h.

\section{Synthetic recover-a-known}
We validate the pipeline on a synthetic flyby --- a rotation-modulated flux series with an injected
period (\vgSynInjected~h), noise, and data gaps --- searched in the \emph{same} wide 14--20~h window
used on the real data. The Lomb--Scargle estimator recovers \vgSynRecovered~h (\vgSynRecoveredOK) and
the Rayleigh fold recovers \vgSynRayPeriod~h (\vgSynRayOK), both inside their bootstrap bands. This is
the control that makes the real-data null interpretable: the method finds a clean rotation when one is
present, so a failure on the real data is a property of the data, not the pipeline.

\section{Results}
\label{sec:results}
On the real encounter volumes the blind periodogram recovers \emph{neither} rotation period
(Table~\ref{tab:periods}).

\textbf{Uranus.} The peak sits at \vgRealUPeriod~h, offset $+1.2$~h from the historical $17.24$~h, and
is \emph{not} band-robust: across four sub-bands it wanders with a spread of \vgRealUSpread~h. It is
not a recovery (\vgRealURecovers); the historical and Hubble values fall inside the $\pm\vgRealUUnc$~h
band only because that band is nearly two hours wide, a vacuous consistency.

\textbf{Neptune.} The peak rails to the upper search bound (\vgRealNPeriod~h), offset far from
$16.11$~h --- the signature of a periodogram dominated by the red-noise / flyby-envelope continuum
rather than a rotational line. It is neither recovered (\vgRealNRecovers) nor consistent
(\vgRealNHistOK). We note explicitly that a narrower search window \emph{can} place the Neptune peak
near $16$~h, but that only demonstrates the result's window-dependence: with an un-tuned wide window
there is no stable rotational peak.

\textbf{Interpretation.} The blind total-power periodogram does not recover the ice-giant rotation
periods, while recovering a clean injected signal in the identical analysis. The auroral total-power
flux is therefore not a clean rotational sinusoid over these flybys; the historical determinations'
beaming and magnetic-longitude modelling --- which a blind periodogram omits --- was essential.
Independently, the rotation-block uncertainties ($\sim$\vgRealUUnc~h) are hundreds of times coarser
than the 28~s by which \citet{lamy2025} refined Uranus, so even a successful blind extraction could
never approach modern precision. The secondary Rayleigh burst fold (Uranus \vgRealURayPeriod~h,
Neptune \vgRealNRayPeriod~h) is likewise uninformative, as expected for a discrete-epoch method on
this emission.

\begin{deluxetable}{lccc}
\tablecaption{Rotation periods (hours)\label{tab:periods}}
\tablehead{\colhead{Source} & \colhead{Uranus} & \colhead{Neptune} & \colhead{method}}
\startdata
Voyager (historical)   & $17.24\pm0.01$ & $16.11$ & radio (modelled) \\
\citet{lamy2025}       & $\vgRealULamy$ & \nodata & HST UV \\
This work (blind)      & $\vgRealUPeriod\pm\vgRealUUnc$ & $\vgRealNPeriod\pm\vgRealNUnc$ & radio (blind LS) \\
recovers historical?   & \vgRealURecovers & \vgRealNRecovers & \\
\enddata
\tablecomments{This work: blind Lomb--Scargle of the PRA flux series over an un-tuned 14--20~h window;
uncertainty = larger of the rotation-block bootstrap $\sigma$ and the sub-band spread. ``Recovered''
requires a band-robust, un-railed peak within one combined-$\sigma$ of the historical value. Neither
planet qualifies.}
\end{deluxetable}

\section{Reproducibility}
The metric, its synthetic recover-a-known, and all unit tests (including a parser test against a
vendored real PRA sample) run offline in CI (\code{python -m jansky\_research.vgpra --offline}). The
real reanalysis downloads the two open PDS-PPI encounter volumes and runs on demand
(\code{scripts/vgpra\_real.py}). Code, tests, and this paper's pipeline are in the
\code{jansky-research} repository, built on the \code{jansky} course library \citep{jansky} and its
\code{frbperiod} periodicity machinery. Draft prepared with the assistance of a large language model
\citep{anthropicclaude}; all numerical results are pipeline-generated and human-verified.

\bibliographystyle{aasjournal}
\bibliography{refs}
\end{document}
