\documentclass[twocolumn]{aastex631}
\input{generated/macros}
\newcommand{\code}[1]{\texttt{#1}}
\shorttitle{A 15-year megaconstellation RFI trend in the e-CALLISTO archive}
\shortauthors{Barbere}
\begin{document}
\title{e-CALLISTO as an Accidental 15-Year RFI Observatory: A Narrowband
Unintended-Emission Trend Across the Megaconstellation Era}
\author[0009-0008-3289-4447]{Joseph Barbere}
\affiliation{Independent researcher}

\begin{abstract}
The e-CALLISTO network of low-cost solar radio spectrometers \citep{benz2009} has recorded
45--870~MHz dynamic spectra worldwide since 2002, making it an unintended long-baseline record of
the radio-frequency-interference (RFI) environment. The only published RFI-trend study on the
archive \citep{perez2020} is a two-epoch (2012 vs.\ 2019) campaign finding a $\sim$2$\times$
interference rise, and it stops in 2019 --- the year Starlink deployment scaled. We mine the
\emph{continuous} archive across the megaconstellation era. From the open FITS record we sample one
representative spectrum per station-month over 2012--2026 at \rfRealNStations\ configuration-stable
stations and reduce each to a burst-immune occupancy metric (a per-channel time median, so a
transient solar burst or satellite pass does not move it). Because e-CALLISTO log-power drifts with
uncalibrated station gain --- a common-mode systematic that would fake a trend --- our primary
metric is a \emph{self-normalizing} one: the excess level at the intrinsic narrowband Starlink
unintended-emission (UEM) lines (125/135/150/175~MHz; \citealt{divruno2023}) over their immediately
adjacent channels, which cancels common-mode station gain and, unlike a fixed FM control, survives
the per-station RFI-avoidance notches that leave many e-CALLISTO instruments not observing the FM
band at all. We attribute the trend to the public Starlink on-orbit count \citep{planet4589}.
\textbf{The result is a systematics-limited null with one intriguing single-station rise.} Across
\rfRealNMonths\ usable station-months, the two stations that sample UEM lines \emph{disagree in
sign}: HUMAIN's line excess (\rfRealHUMAINLines~MHz) rises steeply and significantly
($\rfRealHUMAINSlope$ per year, $p=\rfRealHUMAINP$, correlating with the Starlink count at
$r=\rfRealHUMAINStarCorr$), while ALMATY's (\rfRealALMATYLines~MHz) falls significantly
($\rfRealALMATYSlope$ per year, $p=\rfRealALMATYP$). A genuine \emph{global} megaconstellation
signal should raise every station's UEM lines together, so the sign disagreement (\rfRealNRising\
rising, \rfRealNFalling\ falling; \code{coherent\_rise}=\rfRealCoherent) indicates per-station and
local-RFI effects dominate over any common Starlink signal, and the weakly positive pooled slope
(\rfRealLineExcessSlope\ per year, pooled Starlink correlation only $r=\rfRealCorr$) is a
sample-size artifact of whichever station dominates. We therefore claim no Starlink attribution.
The validated burst-immune metric, the notch-robust self-normalizing line test, the
pipeline-generated cross-station coherence verdict, and the honest per-station systematics
accounting --- including HUMAIN as a flagged candidate for targeted follow-up --- are the
contribution.
\end{abstract}

\keywords{Radio frequency interference (2043) --- Solar radio emission (1522) --- Artificial
satellites (68) --- Astronomy data analysis (1858)}

\section{Introduction}
Large low-Earth-orbit satellite constellations radiate unintended electromagnetic emission (UEM)
that falls inside protected and scientifically used radio bands. Dedicated-instrument campaigns
have detected Starlink UEM directly: LOFAR sees broadband emission between 110 and 188~MHz plus
narrowband features \citep{divruno2023}, and the second-generation satellites are brighter still
\citep{bassa2024}. These are targeted observations with calibrated interferometers. What has not
been examined is whether the \emph{archival} record of a distributed, uncalibrated spectrograph
network already carries the constellation-growth signal --- a question of spectrum management as
much as astronomy, and one the e-CALLISTO archive is uniquely positioned to answer because it spans
the entire deployment era with the same simple receivers.

The only RFI-trend precedent on this archive \citep{perez2020} compares two epochs, 2012 and 2019,
at Spanish sites and reports a $\sim$2$\times$ rise in interference occupancy across the band --- but
it ends in 2019, the year the first operational Starlink batches launched. The 2012--2026
e-CALLISTO record is therefore an unexamined 15-year RFI observatory covering the megaconstellation
ramp, and extending the two-epoch result into a continuous trend, attributed to the public
constellation count, is unclaimed. We do that here, with the honest caveat stated up front: the
archive can ask whether a \emph{post-2019, UEM-band-specific} rise tracks Starlink growth, but it
cannot by itself prove causation, because general terrestrial RFI grows over the same period.

\section{Method}
\label{sec:method}
\textbf{The systematics problem.} e-CALLISTO values are uncalibrated digital log-power. Two effects
would manufacture or mask a trend: (i) \emph{station gain drift} --- receiver and antenna changes
move the whole 45--870~MHz spectrum up or down over years; and (ii) \emph{solar bursts} ---
broadband transients that raise all frequencies for minutes. Both are \emph{common-mode}: they move
the band together. Our metrics are built to cancel both.

\textbf{Burst immunity.} For a band we compute, per channel, the \emph{time median} over the
15-minute sweep --- the persistent occupied level --- and take the median of those channel floors
as the band occupancy. A solar burst or satellite pass occupies a small fraction of the sweep, so
it does not move a time median; the metric reads the persistent RFI floor, not transients. We
verify on synthetic stacks that turning bursts on in up to 90\% of months does not change the
recovered trend (\S\ref{sec:synth}).

\textbf{Gain cancellation, and why the control is not FM.} The natural gain-cancelling metric is a
\emph{differential}: the UEM-band occupancy minus that of a fixed-transmitter control band at the
same station, so a common-mode gain drift subtracts out. The cleanest such control is the FM
broadcast band (88--108~MHz), a strong stable terrestrial transmitter. On real e-CALLISTO data,
however, this fails at the archive's best UEM-band station: HUMAIN (focus code~59) \emph{notches out
the FM band entirely} --- a hardware gap from 84 to 112~MHz with zero sampled channels --- and
notches other sub-bands too. e-CALLISTO operators configure station-specific RFI-avoidance notches,
so a fixed FM control is not universally observed. We therefore make the differential's control
\emph{station-adaptive} (falling back to a sampled clean band below or above the UEM window), and
promote to primary a metric that needs no external control at all.

\textbf{Primary metric: the narrowband line-vs-adjacent excess.} \citet{divruno2023} identify
narrowband Starlink UEM features \emph{intrinsic} to the satellites at 125, 135, 150 and 175~MHz (a
fifth feature at 143.05~MHz is reflected GRAVES space-surveillance radar, not intrinsic, and we
exclude it). A genuine intrinsic UEM line sits \emph{above} the channels immediately around it,
whereas station gain or broadband terrestrial RFI raise a line and its neighbourhood together. For
each sampled UEM line we take the level within $\pm$1~MHz minus the level in the flanking channels
and average. This is self-normalizing \emph{within} the UEM band: it cancels \emph{common-mode}
station gain by construction and is robust to the per-station notches (a line simply drops out if
its region is not sampled). It does \emph{not} cancel differential local contamination of a line's
flanking channels --- which, as \S\ref{sec:real} shows, is exactly the residual systematic that
limits the study. Configuration stability is enforced by requiring a station to keep the \emph{same}
set of sampled UEM lines across its retained months; months with a different sampled grid are
dropped so the trend is not an instrument-reconfiguration artifact. This screen keys on the sampled
line-set only, not on level continuity, so within-set receiver reconfigurations can survive it --- a
limitation the per-station results make visible rather than hide.

\textbf{Attribution.} We regress the metric against decimal year (robust Theil--Sen slope with a
Kendall-$\tau$ significance, so a single anomalous month cannot swing it) and correlate it with the
public cumulative on-orbit Starlink count \citep{planet4589}, which is zero before 2019.4 and rises
monotonically thereafter --- no Space-Track credentials required. As a \emph{qualitative} pre-vs-post
baseline check we also compare the raw UEM occupancy in 2012--2013 against 2018--2019, in the spirit
of the two-epoch \citet{perez2020} campaign --- though this is the gain-contaminated raw level (not
their occupancy statistic) at a different station, so it is a consistency note, not a reproduction.

\section{Synthetic recover-a-known}
\label{sec:synth}
We build \rfSynNMonths\ synthetic monthly spectra with a known UEM-band level rise \emph{and}
narrowband line excesses at the intrinsic UEM frequencies, both shaped like the Starlink count, plus
per-month common-mode gain drift and broadband solar bursts in a configurable fraction of months.
Both metrics recover the injected trend: the secondary differential at \rfSynSlope\ per year
($p=\rfSynTrendP$, $r=\rfSynCorr$ with the injected Starlink shape), and the \emph{primary}
line-vs-adjacent excess at \rfSynLineExcessSlope\ per year ($r=\rfSynLineCorr$) --- so the metric
that headlines the real analysis is validated end-to-end, not only the cross-check. A null control
built from two bands both outside the UEM window does \emph{not} trend (slope \rfSynCtrlSlope\ per
year), confirming the metric does not manufacture a trend from gain drift. Burst immunity is
explicit: the recovered slope is unchanged whether 0\% or 90\% of months carry a broadband burst.
These validations run offline in continuous integration.

\section{The real archive: 2012--2026}
\label{sec:real}
We sampled one representative spectrum per station-month over 2012--2026 at three e-CALLISTO
stations spanning the era (\rfRealNMonths\ station-months after config-stability screening),
reducing each to the burst-immune line-excess and station-adaptive differential. The three stations
sample the UEM band very differently, which is itself the story (Table~\ref{tab:stations}). Only
two of the three observe UEM lines with a stable core-plus-flank grid; GLASGOW samples no UEM line
region stably and drops out of the line analysis entirely.

\begin{deluxetable}{lccccc}
\tablecaption{Per-station UEM-line-excess trend, 2012--2026\label{tab:stations}}
\tablehead{\colhead{Station} & \colhead{UEM lines (MHz)} & \colhead{months} &
\colhead{slope/yr} & \colhead{$p$} & \colhead{$r$(Starlink)}}
\startdata
HUMAIN  & \rfRealHUMAINLines  & \rfRealHUMAINNMonths  & $\rfRealHUMAINSlope$  & \rfRealHUMAINP  & $\rfRealHUMAINStarCorr$ \\
ALMATY  & \rfRealALMATYLines  & \rfRealALMATYNMonths  & $\rfRealALMATYSlope$  & \rfRealALMATYP  & $\rfRealALMATYStarCorr$ \\
GLASGOW & \rfRealGLASGOWLines & \rfRealGLASGOWNMonths & \nodata               & \nodata         & \nodata \\
\enddata
\tablecomments{The line-excess is the level at the sampled Starlink UEM lines over their adjacent
channels, robustly trended (Theil--Sen slope, Kendall-$\tau$ $p$) and correlated with the public
Starlink on-orbit count. GLASGOW samples no UEM line with a stable core-plus-flank grid.}
\end{deluxetable}

\textbf{The two line-sampling stations disagree in sign.} HUMAIN's line excess (lines
\rfRealHUMAINLines~MHz) rises strongly and significantly ($\rfRealHUMAINSlope$ per year,
$p=\rfRealHUMAINP$), positively correlated with the Starlink count ($r=\rfRealHUMAINStarCorr$), and
its raw UEM occupancy rose by \rfRealHUMAINPerez\ log-units between 2012--2013 and 2018--2019 ---
qualitatively consistent with the pre-Starlink rise that \citet{perez2020} found in their two-epoch
campaign, though this is the gain-contaminated raw level at a different station, so we read it as a
consistency note rather than a reproduction. Taken alone, HUMAIN looks like a textbook
megaconstellation UEM signal. But ALMATY --- overhead the same global constellation --- trends the
\emph{opposite} way: its line excess (lines \rfRealALMATYLines~MHz) \emph{falls} significantly
($\rfRealALMATYSlope$ per year, $p=\rfRealALMATYP$, $r=\rfRealALMATYStarCorr$).

\textbf{The cross-station coherence test is the verdict.} A real global Starlink UEM signal should
raise every station's UEM lines together. Our pipeline computes, from the significant per-station
slopes, whether they agree in sign: they do not (\rfRealNRising\ rising, \rfRealNFalling\ falling;
\code{coherent\_rise}=\rfRealCoherent). The weakly positive network-pooled slope
(\rfRealLineExcessSlope\ per year, $p=\rfRealTrendP$) with only a $r=\rfRealCorr$ Starlink
correlation is therefore a sample-size artifact of the more populous rising station, not a coherent
trend. \textbf{We claim no Starlink attribution and no megaconstellation detection.} The result is
consistent with per-station effects limiting any common signal: local RFI-environment change,
receiver and antenna reconfiguration (e-CALLISTO instruments are uncalibrated and operator-tuned),
and station-specific occupancy of the flanking channels that normalize each line. This is exactly
the systematics limit the differential/multi-station design was built to expose.

\textbf{A caveat on the coherence test itself.} With only two line-sampling stations (GLASGOW drops
out), the test is an underpowered global-vs-local discriminator, and opposite signs do not
\emph{strictly} rule out a real global signal: the two stations sample partly different lines
(Table~\ref{tab:stations}), and because the line-excess is self-normalized against its flanks, a
station whose flank channels accrue local terrestrial RFI faster than the line can register a
\emph{falling} excess even under a rising true signal. So the test shows the data are dominated by
per-station systematics; it does not prove the absence of any Starlink contribution. Because our
conclusion is a null, this asymmetry is conservative --- it can only make us \emph{less} likely to
claim a detection.

\textbf{HUMAIN as a flagged candidate.} The honest exception is that HUMAIN's rise is real,
significant, and Starlink-correlated in the right band. We cannot attribute it to Starlink from
spectral archive data alone --- a co-located general-RFI rise would look identical --- but it is the
right target for a follow-up that the archive cannot provide: satellite-pass-gated occupancy (does
the excess appear only when Starlink satellites are above the horizon at HUMAIN?), which would break
the local-RFI degeneracy. We flag it, and claim nothing more.

\section{Reproducibility}
The metric, its synthetic recover-a-known, and all unit tests run offline in CI
(\code{python -m jansky\_research.rfitrend --offline}). The real census streams the open e-CALLISTO
FITS archive one small gzipped file at a time, entirely in memory
(\code{scripts/rfitrend\_real.py}). Code, tests, and this paper's pipeline are in the
\code{jansky-research} repository, built on the \code{jansky} course library \citep{jansky}. Draft
prepared with the assistance of a large language model \citep{anthropicclaude}; all results are
pipeline-generated and human-verified.

\bibliographystyle{aasjournal}
\bibliography{refs}
\end{document}
