\documentclass[twocolumn]{aastex631}
\input{generated/macros}
\newcommand{\code}[1]{\texttt{#1}}
\shorttitle{Cassini SKR proximity duty-cycle census}
\shortauthors{Barbere}
\begin{document}
\title{Is SKR Detection Occurrence a Saturn-Proximity Law? A Sensitivity-Null
Census from Cassini/RPWS, Anchored to the Known Rotation Period}
\author[0009-0008-3289-4447]{Joseph Barbere}
\affiliation{Independent researcher}

\begin{abstract}
Saturn Kilometric Radiation (SKR) is Saturn's dominant auroral radio emission, and its dual
rotation-modulation periods ($\sim$10.6/10.8~h north/south) are tracked end-to-end across the
Cassini mission \citep{gurnett2009,fischer2015,gurnett2016,provan2019}. We port the occurrence-census
method of a sibling Jovian-decametric study to Saturn: over \skrNBins\ one-minute RPWS
key-parameter bins from the 2017 proximal (Grand Finale) orbits, we band-integrate the electric
spectral density over the SKR band, flag SKR-active bins by a background$+k\sigma$ threshold, and
fold detection occurrence against the Cassini--Saturn range from JPL Horizons. Two results, in
order of confidence. First, as a pipeline \emph{validation}, a Lomb-Scargle analysis of the
intensity series recovers the dual SKR rotation period at \skrRealAnchorPeriod\ and
\skrRealPeriodTwo~h --- within \skrRealAnchorDev\% of the published late-mission values
\citep{provan2019}, so the flux series and detection are sound; we claim no new period result, the
record being already complete. Second, the occurrence census: the SKR-active duty cycle rises from
\skrRealDutyFar\% at \skrRealRangeFar~$R_S$ to \skrRealDutyNear\% at \skrRealRangeNear~$R_S$, a raw
near/far ratio of \skrRealNearFar. But this is \emph{almost entirely a $1/r^2$ sensitivity effect}:
correcting each bin's flux to a common range and re-thresholding collapses the ratio to
\skrRealSensCorrNearFar. Even that residual is not clean --- the near and far range bins differ in
sub-spacecraft latitude by \skrRealAbsLatSpan$^\circ$, so SKR's known latitude-dependent visibility
is entangled with range and is not removed. The honest result is therefore a near-null: the
apparent SKR ``proximity law'' is dominated by inverse-square detection sensitivity, with at most a
$\sim$\skrRealSensCorrNearFar$\times$ residual that we cannot cleanly separate from viewing
geometry. The value is the bounded, honestly-framed measurement and the validated, reusable
tooling --- not a proximity law.
\end{abstract}

\keywords{Saturn (1426) --- Planetary magnetospheres (997) --- Radio astronomy (1338) ---
Non-thermal radiation sources (1119) --- Auroral kilometric radiation}

\section{Introduction}
Saturn Kilometric Radiation (SKR; $\sim$100--400~kHz, cyclotron-maser emission from the auroral
regions) is modulated near Saturn's rotation period, with the striking property of \emph{two}
close periods from the northern and southern auroral systems that drift and swap across equinox
\citep{gurnett2009,fischer2015}. That periodicity record is complete: Fischer et al.\ tracked it
through early 2013 \citep{fischer2015}, and Gurnett et al.\ and Provan et al.\ carried it to the
end of the mission \citep{gurnett2016,provan2019}. What has \emph{not} been asked is a simpler
occurrence question, by analogy to a proximity--occurrence law reported for Jovian decametric
emission from Juno: does SKR \emph{detection} occurrence depend on the observer's distance from
the planet? Cassini's range to Saturn varied by more than an order of magnitude across each orbit,
so its own key-parameter archive answers the question directly. The occurrence and visibility of
SKR versus observer local time and latitude have been mapped \citep{lamy2008,ye2016}; the
observer-\emph{range} axis, and specifically whether a $1/r^2$ sensitivity trend leaves any
residual occurrence dependence, has not. We build that census, using the known rotation period as
a built-in validation.

\section{Data and method}
\label{sec:method}
We use the Cassini/RPWS 60-second summary key parameters
(\code{CO-V/E/J/S/SS-RPWS-4-SUMM-KEY60S-V1.0}; \citealt{rpws_key60s}), which provide per-minute
calibrated electric spectral densities on a 73-channel electric-antenna grid but \emph{no}
pre-integrated SKR flux --- so we integrate the electric spectral density over the SKR band
(0.1--1.2~MHz) ourselves (\code{read\_key\_params} $+$ \code{band\_integrated\_flux}, tested
against the fixed-width PDS3 record layout). A bin is SKR-active when its band flux exceeds a
robust background ($25$th log-flux percentile) plus $k\sigma$ (the sibling census's
detection rule). The Cassini--Saturn range and sub-spacecraft latitude come from JPL Horizons
(target Saturn, observer Cassini). We analyze the 2017 proximal orbits (\skrNBins\ one-minute
bins), where the range sweeps repeatedly over $\sim$8--21~$R_S$.

\textbf{Validation.} Before any occurrence claim we run a Lomb-Scargle periodogram of the
log-intensity series in the physically-motivated Saturn-rotation band (10.4--11.0~h, bracketing
the established $\sim$10.6--10.8~h periods). It must recover the known period; it does
(\S\ref{sec:results}). A broader search additionally shows a stronger, unexplained sub-11~h
feature near $\sim$10.34~h --- plausibly a sampling/aliasing artefact of the orbit-modulated
duty cycle --- which we deliberately exclude from the rotation anchor; stated, not hidden.

\textbf{Proximity census and its null model.} We bin the SKR-active flag by Cassini--Saturn range
quartile and report the duty cycle per bin. The dominant confound is inverse-square sensitivity:
at smaller range the \emph{same} intrinsic emission produces a larger flux and clears the fixed
threshold more often. We test it directly --- correct each bin's flux to a common reference range
($S\to S\,(r/r_\mathrm{ref})^2$), re-threshold, and recompute the near/far ratio. If the trend is
pure sensitivity the corrected ratio is unity; any excess is the intrinsic$+$beaming part. A second
confound is latitude: SKR is auroral and beamed, and range correlates with sub-spacecraft latitude
along the orbit, so we report the $|\mathrm{lat}|$ span between the range bins to make the residual
entanglement visible rather than assert it away.

\section{Results}
\label{sec:results}
\textbf{Rotation-period anchor (validation).} The Lomb-Scargle analysis recovers the dual SKR
rotation period at \skrRealAnchorPeriod\ and \skrRealPeriodTwo~h, matching the published
late-mission northern ($\sim$10.79~h) and southern ($\sim$10.68~h) values \citep{provan2019} to
better than 1\%. This is a reproduction, not a new result --- its role is to certify that the
band-integrated flux series and the detection are physically faithful before we trust the
occurrence census.

\textbf{Occurrence census (a near-null).} Across \skrNBins\ bins the overall SKR-active duty cycle
is \skrRealDutyPct\%. The raw duty cycle rises monotonically with proximity, from \skrRealDutyFar\%
at $\sim$\skrRealRangeFar~$R_S$ to \skrRealDutyNear\% at $\sim$\skrRealRangeNear~$R_S$ (near/far
\skrRealNearFar). The $1/r^2$ null model accounts for almost all of it: after correcting the flux
to a common range and re-thresholding, the near/far ratio falls to \skrRealSensCorrNearFar. So the
apparent proximity ``law'' is dominated by inverse-square detection sensitivity. The residual
$\sim$\skrRealSensCorrNearFar$\times$ is moreover not cleanly intrinsic: the near and far range
quartiles differ in $|$sub-spacecraft latitude$|$ by \skrRealAbsLatSpan$^\circ$, and SKR
visibility is latitude-dependent \citep{lamy2008,ye2016}, so this residual is entangled with
viewing geometry and is at most an upper bound on any intrinsic occurrence-versus-range effect.

\textbf{Honest interpretation.} The result is a bounded near-null, not a proximity law: SKR
detection occurrence versus Cassini range is explained by $1/r^2$ sensitivity to within a
$\sim$\skrRealSensCorrNearFar$\times$ residual that we cannot separate from the
\skrRealAbsLatSpan$^\circ$ latitude difference between bins. This is a cleaner statement than the
sibling Jovian-decametric census, whose $\sim$180$\times$ raw proximity ratio was reported as
proximity-dominated \emph{detection} but without an explicit sensitivity null --- here the null is
computed and it removes essentially the whole effect.

\section{Conclusions}
Cassini's own RPWS key-parameter archive yields a bounded near-null: the raw $\sim$3$\times$ rise
in SKR detection occurrence from $\sim$21 to $\sim$8~$R_S$ across the 2017 proximal orbits is
almost entirely $1/r^2$ detection sensitivity --- the null model collapses it to
$\sim$\skrRealSensCorrNearFar$\times$, and that residual is itself entangled with a
\skrRealAbsLatSpan$^\circ$ latitude difference between the range bins. The one firm result is the
validation: the pipeline recovers the published dual rotation period to \skrRealAnchorDev\%. Two
follow-ons use the same tooling: extending the range baseline with earlier, higher-apoapsis orbits
(to $\sim$60~$R_S$), where a real intrinsic dependence would have more leverage against the
sensitivity floor; and a proper SKR beaming/visibility model to disentangle the residual from
latitude. The sensitivity-null method, the rotation-period anchor, and the tested KEY60S tooling
are the deliverable.

\textit{Reproducibility.} \code{jansky\_research.skr} (parser $+$ detection $+$ Lomb-Scargle $+$
proximity census, synthetic recover-a-known in core CI) $+$ \code{scripts/skr\_real.py} (the
PDS-PPI KEY60S downloader); every number above is pipeline-generated
(\code{results/skr\_metrics.json} $\to$ macros). Data: \citet{rpws_key60s}; geometry from JPL
Horizons.

\bibliography{refs}
\bibliographystyle{aasjournal}
\end{document}
