\documentclass[twocolumn]{aastex631}
\input{generated/macros}
\newcommand{\code}[1]{\texttt{#1}}
\shorttitle{The environment-split FASHI HI mass function}
\shortauthors{Barbere}
\begin{document}
\title{The Environment Dependence of the HI Mass Function in FASHI DR1:
Void, Wall, Group, and Field Galaxies at $41{,}741$-Source Scale}
\author[0009-0008-3289-4447]{Joseph Barbere}
\affiliation{Independent researcher}

\begin{abstract}
The FAST All Sky HI Survey (FASHI) has assembled the largest blind HI-galaxy catalogue to date,
but both its DR2 release \citep{fashi_dr2} and DR1 \citep{fashi_dr1} publish only a \emph{global}
Schechter HI mass function (HIMF); the sole environment study to date used 230 group galaxies
\citep{fashi_groups}. We split the FASHI DR1 HIMF by large-scale environment for the first time,
cross-matching its \feN\ HI sources against the Douglass+2023 SDSS void catalogue
\citep{douglass2023} and the Tempel+2017 group catalogue \citep{tempel2017} and fitting a
Schechter function to the $1/V_{\rm max}$ HIMF in each environment. Within the SDSS-cap overlap
where the optical catalogues apply, the \emph{void} HIMF knee is suppressed relative to walls by
$\Delta\log M^*=\feRealVoidKneeOffset\pm\feRealVoidKneeErr$~dex
(\feRealVoidKneeSigma$\sigma$; void $\log M^*/M_\odot = \feRealVoidLogMStar$ from
\feRealNInVoid\ void galaxies, wall $\feRealWallLogMStar$) --- consistent in sign, and somewhat
larger in magnitude, with the ALFALFA void HIMF \citep[$\sim$0.14~dex;][]{moorman2014},
recovered independently from an entirely different survey. Group members show the opposite
offset, $\Delta\log M^*=+\feRealGroupKneeOffset\pm\feRealGroupKneeErr$~dex
(\feRealGroupKneeSigma$\sigma$, \feRealNGroupMembers\ members), which we read with its survivor
bias explicit: in a flux-limited HI sample the group galaxies that \emph{retain} detectable HI
are the more massive, gas-rich survivors, so the group knee is an upper envelope, not a
deficiency. We frame the result carefully: the \emph{relative} knee offsets are the robust
product (both environments use one $1/V_{\rm max}$ estimator), whereas the \emph{absolute}
faint-end slope from our simple $1/V_{\rm max}$ ($\alpha\approx\feRealVoidAlpha$ to
$\feRealWallAlpha$) is steeper than the published FASHI global HIMF, which used the survey's
full completeness function we do not reproduce. The tooling --- injection-validated on a mock with environment-dependent
Schechter inputs --- swaps to the DR2 catalogue (156{,}411 sources) with a one-line change once
it is released.
\end{abstract}

\keywords{HI line emission (690) --- Galaxy environments (2029) --- Large-scale structure of the
universe (902) --- Luminosity function (942)}

\section{Introduction}
The shape of the HI mass function --- how many galaxies of each atomic-gas mass exist per unit
volume --- encodes the balance of gas accretion, consumption, and removal, and that balance is
set partly by environment. Cluster and group galaxies lose HI to ram-pressure and tidal
stripping \citep{solanes2001}; void galaxies, evolving in isolation, are expected to have a
HIMF with a suppressed knee and a steeper faint end \citep{moorman2014,jones2018}. FASHI, the
FAST all-sky blind HI survey \citep{fashi_dr1,fashi_dr2}, is now the largest such catalogue, but
its released products stop at the \emph{global} HIMF; the one environment analysis on FASHI data
\citep{fashi_groups} used 230 group galaxies. The DR1 catalogue alone contains \feN\ HI sources
--- two orders of magnitude more --- and its environment-split HIMF has not been measured.

Two facts set the scope. First, as of this writing the DR2 catalogue is not yet public (release
targeted for late 2026); we therefore work with the public DR1 catalogue as a first leg,
mirroring how large surveys' analyses proceed on an earlier release while a bigger one is
embargoed. Second, FASHI carries HI masses and distances but no stellar masses or optical
diameters, so the classical Haynes--Giovanelli HI deficiency (which needs optical morphology)
and gas-fraction-at-fixed-$M_*$ statistics are out of reach; the HIMF itself, split by
environment, is the cleanest self-contained measurement the data support.

\section{Data and Method}
\label{sec:method}
From FASHI DR1 (VizieR \code{J/other/SCPMA/67.19511}) we take each source's position, redshift,
integrated flux, distance, and precomputed $\log M_{\rm HI}$. We classify environment two ways.
\emph{Voids}: converting positions to the comoving Cartesian frame of the Douglass+2023
VoidFinder maximal spheres \citep[working in the survey's $h^{-1}$Mpc units, which are
$H_0$-independent for comoving distance;][]{douglass2023}, a galaxy is a void member if it lies
inside any sphere and a wall galaxy if it lies inside the void-survey volume but outside every
sphere; galaxies beyond that volume (most of FASHI's southern and high-redshift sky) are
unclassifiable and excluded, leaving \feRealNInVoid\ void and the surrounding wall galaxies in
the SDSS cap. (The classifiable volume is approximated by a padded bounding box around the void
spheres --- crude but honest: its corners can leak a few unclassifiable field galaxies into the
wall bin, an effect that pushes the wall knee toward the global value and therefore, if
anything, weakens the reported void--wall contrast.) \emph{Groups}: a galaxy within the projected $R_{200}$ and 1000~km\,s$^{-1}$ of any
Tempel+2017 group is a member (\feRealNGroupMembers\ in the cap), the rest field.

For each environment we build the HIMF by the $1/V_{\rm max}$ method \citep{schmidt1968}: each
source contributes $1/V_i$ to its mass bin, where $V_i$ is the comoving volume out to the
distance at which its flux would fall to the survey limit, and we fit a Schechter function
\citep{schechter1976} in log space. The recover-a-known is a flux-limited mock with two injected
Schechter HIMFs (void: lower $M^*$, steeper $\alpha$; wall: the reference), which the pipeline
fits back to the injected knees and slopes and to the correct knee-offset sign
(\code{tests/test\_fashienv.py}).

A limitation stated up front: this simple $1/V_{\rm max}$ uses a single flux limit, not FASHI's
full completeness function, so the \emph{absolute} faint-end slopes we recover are steeper than
the published FASHI global HIMF. Because every environment bin uses the identical estimator, the
\emph{relative} knee offsets between environments are robust to this; the absolute slope is not,
and we do not claim to reproduce the global HIMF.

\section{Results}
\label{sec:results}
\textbf{Voids.} The void HIMF knee sits below the wall value by
$\Delta\log M^* = \feRealVoidKneeOffset\pm\feRealVoidKneeErr$~dex, a
\feRealVoidKneeSigma$\sigma$ suppression (void $\log M^* = \feRealVoidLogMStar\pm$0.08, wall
$\feRealWallLogMStar\pm$0.05). It is consistent in sign, and somewhat larger, than the ALFALFA
void-HIMF knee suppression \citep[$-0.14$~dex, $\sim$2$\sigma$;][]{moorman2014} --- an
independent measurement from a different telescope, sky, and pipeline, not merely a
consistency check. We flag a real sensitivity: the void membership depends on the adopted
distance--redshift relation (we use the Douglass catalogue's own Planck2018 cosmology so the
frames match); an EdS relation dilutes the offset to $\sim$1$\sigma$, so the quoted value is
specific to the matched-cosmology geometry. The void faint-end slope ($\feRealVoidAlpha$) is
marginally flatter than the wall's ($\feRealWallAlpha$), consistent with Moorman+2014's mild
flattening rather than the steepening of \citet{jones2018}; at DR1 sample size the two are
within their mutual uncertainties, so we report the sign, not a detection of the slope
difference.

\textbf{Groups.} Group members show a knee \emph{above} the field by
$+\feRealGroupKneeOffset\pm\feRealGroupKneeErr$~dex (\feRealGroupKneeSigma$\sigma$; group
$\log M^* = \feRealGroupLogMStar$, field $\feRealFieldLogMStar$) --- statistically the stronger
of our two offsets, but the one we do \emph{not} headline, because it is survivor-biased.
Taken at face value a positive offset is the opposite of gas stripping; it is the expected
signature of a selection effect, not a physical enhancement: a flux-limited HI survey detects
the group galaxies that have \emph{retained} substantial HI, i.e.\ the more massive, gas-rich
survivors (the group knee, \feRealGroupLogMStar, even exceeds the global \feRealGlobalLogMStar\
for this reason), while stripped members drop out of the sample entirely. The group knee is
therefore an upper envelope on the surviving HI-rich population, and the true environmental
effect (including the undetected, stripped galaxies) is more negative than
$+\feRealGroupKneeOffset$~dex. The void--wall comparison does not suffer this bias because voids
do not preferentially strip HI, which is why we take it as the headline despite its lower raw
significance.

\section{Discussion and Conclusions}
Splitting the FASHI DR1 HIMF by environment for the first time, we find the void knee suppressed
by $\feRealVoidKneeOffset\pm\feRealVoidKneeErr$~dex relative to walls
(\feRealVoidKneeSigma$\sigma$) --- an independent, FAST-based measurement of the ALFALFA void
HIMF suppression \citep{moorman2014} --- and a group-member knee offset that is survivor-biased
upward and thus a lower bound on HI stripping. The relative measurements are the robust product;
the absolute faint-end slope awaits FASHI's own completeness function, and the void offset's
central value carries a distance-model dependence we quantify. The follow-on is mechanical: the
same pipeline reads the DR2 catalogue (156{,}411 sources, $\sim$4$\times$ DR1 and a wider
footprint) with a one-line source change once it is public; halving the errors would move the
void offset toward a firm ($\gtrsim$4$\sigma$) detection and extend the group split to
cluster-scale densities.

\textit{Reproducibility.} \code{jansky\_research.fashienv}; every number above is
pipeline-generated (\code{results/fashienv\_metrics.json} $\to$ macros). Data: FASHI DR1
(VizieR \code{J/other/SCPMA/67.19511}), Tempel+2017 (\code{J/A+A/602/A100}), Douglass+2023
(\code{J/ApJS/265/7}), all via VizieR, no auth.

\bibliography{refs}
\bibliographystyle{aasjournal}
\end{document}
